---
title: "MN College Information"
output: 
  flexdashboard::flex_dashboard:
    theme:
      version: 4
      bootswatch: flatly
runtime: shiny
---

About
==========================

Column {data-width = 650}
---------------------------

#### Welcome!

**Hello high-schoolers! As recent high school graduates, we understand that navigating your college decision is difficult. We hope you find this website to be useful as you investigate Minnesota colleges and universities. We have conducted an analysis of multiple factors you might be considering, but please note that everyone cares about different things when evaluating colleges. This is just one place to start! Good luck!**

**-Solveig, Ela, and Morgan**


#### Information

**Authors:**

Solveig Senf, Ela Kanade, & Morgan Ryan

**Data & College Resources**

The data used in this project is from National Center for Education Statistics' IPEDS data base. We compiled data from multiple data sets which all can be accessed [here](https://nces.ed.gov/ipeds/datacenter/DataFiles.aspx?gotoReportId=7&fromIpeds=true&sid=3ce8c2b7-5a48-42a0-baf2-9089ef924924&rtid=7). 

Additionally, we have taken data from ["Niche"](https://www.niche.com/?ref=colleges). It is an excellent resource for information on colleges. Just search for a school!

If you still have unanswered questions about the colleges you are interested in, we recommend visiting [U.S. News Best Colleges](https://www.usnews.com/best-colleges) or spending some time on the college's own website.  

Column
---------------------------

```{r, echo = FALSE}
renderImage({
    list(src = "images/graduate.png", width = 600, height = 600)
  })
```

Minnesota College Map
==========================

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(stringr)
library(rvest)
library(polite)
library(sf)
library(maps)
library(viridis)
library(leaflet)
library(htmltools)
library(readr)
library(ggplot2)
library(ggthemes)
library(ggrepel)
library(plotly)
library(kableExtra)
```

```{r, echo = FALSE, include = FALSE}
#Data Cleaning!
ic <- read_csv("data/ic2023.csv") |>
  select(college_id = UNITID, 
         study_abroad = SLO6, 
         room_capacity = ROOMCAP, 
         ncaa = ASSOC1, 
         college_ran_housing = ROOM, 
         meals_per_week = MEALSWK) 
  
ic_ay <- read_csv("data/ic2023_ay.csv") |>
  select(college_id = UNITID,
         in_state_tuition = TUITION2,
         food_housing_fee = CHG5AY3) 

hd <- read_csv("data/hd2023 2.csv") |>
  select(college_id = UNITID,
         college_name = INSTNM,
         city = CITY,
         state = STABBR,
         open_or_closed = CYACTIVE,
         size_type = INSTSIZE,
         longitude = LONGITUD,
         latitude = LATITUDE,
         public_private = CONTROL)

effy <- read_csv("data/effy2023.csv") |> 
  select(college_id = UNITID,
         student_type = EFFYLEV,
         total_enrollment = EFYTOTLT,
         total_men = EFYTOTLM,
         total_women = EFYTOTLW) |>
  filter(student_type == "2") #filters for undergraduate enrollment numbers 

s <- read_csv("data/s2023_sis.csv") |>
  select(college_id = UNITID,
         employment_type = FACSTAT, 
         total_faculty = SISTOTL) |>
  filter(employment_type == "10") |> #staff with faculty status
  select(-employment_type)
```

```{r, echo = FALSE, include = FALSE}
#Joining clean data sets
ic_full <- ic |>
  left_join(ic_ay)  

ic_hd <- ic_full |>
  left_join(hd)

college_data <- ic_hd |>
  left_join(effy)

college_data2 <- college_data |>
  left_join(s)

mn_college_data <- college_data2 |>
  filter(state == "MN") |>
  arrange(college_name)
```

Column {data-width = 100}
-------------------------

**Wondering about location?**

Check out this map of Minnesota colleges and universities! Click on each marker for the college name, city, and enrollment!

```{r, echo = FALSE}
renderImage({
    list(src = "images/pointing_grad.png", width = 400, height = 450)
  })
```

Column {data-width = 600}
-------------------------

**Minnesota College Map**

```{r, echo = FALSE}

mn_map <- leaflet() |>
  addTiles() |>
  setView(lng = -94.6859, lat = 46.7296, zoom = 6)


mn_map <- mn_map |>
  addMarkers(data = mn_college_data,
             ~longitude, ~latitude,
             popup = ~paste("<b>Name:</b>", college_name, "<br><b>City:</b>", city, "<br><b>Total Enrollment</b>:", total_enrollment))
mn_map
```


Tuition & Type
==========================

```{r, echo = FALSE, include = FALSE}
#tibble for tuition by size graph
tuition_by_size <- mn_college_data |>
  select(size_type, in_state_tuition, public_private) |>
  filter(!is.na(in_state_tuition), in_state_tuition != ".") |>
  group_by(size_type, public_private) |>
  mutate(in_state_tuition = as.numeric(in_state_tuition),
         size_type = as.factor(size_type),
         public_private = as.factor(public_private),
         size_type = fct_recode(size_type, 
                                "< 1,000" = "1",
                                "1,000 - 4,999" = "2",
                                "5,000 - 9,999" = "3",
                                "10,000 - 19,999" = "4",
                                "20,000 +" = "5"),
         public_private = fct_recode(public_private,
                                     "Public" = "1",
                                     "Private (not-for-profit)" = "2",
                                     "Private (for-profit)" = "3")) |>
  summarize(mean_tuition = mean(in_state_tuition)) 
```

Column {data-height=950}
-------------------------

```{r, echo = FALSE}
ui <- fluidPage(
  checkboxInput("check_private", label = "Show Private Institutions?", 
                choices = c("Public" = "Public",
                "Private (not for profit)" = `Private (not-for-profit)`,
                "Private (for-profit)" = `Private(for-profit)`)),

  tableOutput("college_table")
)

server <- function(input, output) {
  output$college_table <- renderTable({
    if (input$check_private) {
      filtered_mn_college_data |>
        filter(`College Type` != ("Public"|`Private (not-for-profit)`))
    } else {
      filtered_mn_college_data |>
        filter(`College Type` == "Public")
    } else {
      filtered_mn_college_data |>
        filter(`College Type` == ("Public"|`Private (for-profit)`))
    }
    
  })
}

shinyApp(ui = ui, server = server)

```


```{r, echo = FALSE, include = FALSE}
filtered_mn_college_data <- mn_college_data |>
  select(college_name, public_private) |>
  filter(public_private != "NA") |>
  mutate(public_private = as.factor(public_private),
         public_private = fct_recode(public_private,
                                     "Public" = "1",
                                     "Private (not-for-profit)" = "2",
                                     "Private (for-profit)" = "3")) |>
  rename(college_name, "College Name" = "college_name",
         public_private, "College Type" = "public_private") |>
  arrange(`College Type`) 
  #kable() displays table nicely 
```

Column {data-height = 1500}
-------------------------

**Tuition & Size**

```{r, echo = FALSE, warning = FALSE}
inputPanel(
  checkboxInput("type_check", "Color by Institution Type?", value = FALSE)
)
```

```{r, echo = FALSE, warning = FALSE}
renderPlot({
     if (input$type_check) 
        ggplot(tuition_by_size, aes(x = size_type, y = mean_tuition))  + 
          geom_bar(aes(fill = public_private), stat = "identity") +
          theme_classic() +
          scale_fill_colorblind() +
          labs(title = "Average Tuition by Institution Size",
               subtitle = "Colored by Institution Type",
               x = "Size",
               y = "Average Tuition",
               fill = "Type") 
     else if (!input$type_check)      
        ggplot(tuition_by_size, aes(x = size_type, y = mean_tuition))  + 
          geom_bar(stat = "identity") +
          theme_classic() +
          scale_fill_colorblind() +
          labs(title = "Average Tuition by Institution Size",
               subtitle = "Colored by Institution Type",
               x = "Size",
               y = "Average Tuition",
               fill = "Type")
})
```

**Tuition & Type**

```{r, echo = FALSE, warning = FALSE}
mn_college_data |>
  select(in_state_tuition, public_private) |>
  filter(!is.na(in_state_tuition), in_state_tuition != ".") |>
  group_by(public_private) |>
  mutate(in_state_tuition = as.numeric(in_state_tuition),
         public_private = as.factor(public_private),
         public_private = fct_recode(public_private,
                                     "Public" = "1",
                                     "Private (not-for-profit)" = "2",
                                     "Private (for-profit)" = "3")) |>
  summarize(mean_tuition = mean(in_state_tuition)) |>
  ggplot(aes(x = public_private, y = mean_tuition, fill = public_private)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(mean_tuition, 1)), vjust = -0.5, size = 3) +
  labs(title = "Average Tuition by Institution Type",
       x = "Type",
       y = "Average Tuition") +
  theme_classic() +
  scale_fill_colorblind(guide = "none") 
```

Student/Faculty Ratio
==========================

Column {data-height = 1500}
-------------------------

```{r, echo = FALSE, include = FALSE}
#tibble for student-to-faculty interactive plot
interactive_data <- mn_college_data |>
  select(total_enrollment, total_faculty, college_name, public_private) |>
  filter(total_enrollment != "NA",
         total_faculty != "NA") |>
  mutate(student_to_fac = total_enrollment/total_faculty,
         total_enrollment = as.numeric(total_enrollment),
         public_private = as.factor(public_private),
         public_private = fct_recode(public_private,
                                     "Public" = "1",
                                     "Private (not-for-profit)" = "2",
                                     "Private (for-profit)" = "3")) 
```

```{r, echo = FALSE, include = FALSE}
#tibble for student-to-faculty interactive plot w/out outliers
interactive_data_rm <- mn_college_data |>
  select(total_enrollment, total_faculty, college_name, public_private) |>
  filter(total_enrollment != "NA",
         total_faculty != "NA") |>
  mutate(student_to_fac = total_enrollment/total_faculty,
         total_enrollment = as.numeric(total_enrollment),
         public_private = as.factor(public_private),
         public_private = fct_recode(public_private,
                                     "Public" = "1",
                                     "Private (not-for-profit)" = "2",
                                     "Private (for-profit)" = "3")) |>
  filter(college_name != "University of Minnesota-Twin Cities",
         college_name != "Capella University")
```

```{r, echo = FALSE, warning = FALSE}
inputPanel(
  checkboxInput("type_check", "Remove Outliers?", value = FALSE)
)
```

```{r, echo = FALSE, warning = FALSE}
renderPlotly({
     if (!input$type_check) {
        interactive <- ggplot(interactive_data, aes(x = total_enrollment, y = student_to_fac)) +
        geom_point(aes(text = college_name, color = public_private)) +
        geom_text_repel(aes(label = college_name)) +
        labs(title = "Total Enrollment Vs. Student-to-faculty Ratio",
            x = "Enrollment",
            y = "Student-to-faculty",
            color = "Institution Type") +
        theme_classic() +
        scale_color_colorblind() 
  
        ggplotly(interactive)
     }
     else if (input$type_check) {
        interactive2 <- ggplot(interactive_data_rm, aes(x = total_enrollment, y = student_to_fac)) +
        geom_point(aes(text = college_name, color = public_private)) +
        geom_text_repel(aes(label = college_name)) +
        labs(title = "Total Enrollment Vs. Student-to-faculty Ratio",
            x = "Enrollment",
            y = "Student-to-faculty",
            color = "Institution Type") +
        theme_classic() +
        scale_color_colorblind() 
        
        ggplotly(interactive2)
     }
})
```

Majors
==========================


```{r}
#| message: FALSE
#| eval: FALSE

library(httr)
library(httr2)

college_tibble <- tibble(College_name = character(), Major = character())

for(i in 1:5) {
  new <- scrape_majors(college_for_majors[2,])
  college_tibble <- rbind(college_tibble, new)
}

college_tibble

  url <- str_c("https://www.niche.com/colleges/", college_for_majors[1,], "/majors/")
  Sys.sleep(0.5) 
  one_college <- GET(url)
  details <- content(one_college, "parse")
  college_tibble[1,1] <- details$College_name
  college_tibble[1,2] <- details$Major

```

```{r, include = FALSE}
#Check if scraping is allowed
robotstxt::paths_allowed("https://www.niche.com/colleges/search/best-colleges/")

session <- bow("https://www.niche.com/colleges/carleton-college/majors/", force = TRUE)

top_majors <- scrape(session) |>
  html_nodes(".popular-entity__name") |>
  html_text()
top_majors
```

```{r, echo = FALSE}
#A function that takes college names in the correct format (lower case with dashes) 
scrape_majors <- function(college) {
 
    url <- str_c("https://www.niche.com/colleges/", college, "/majors/")
    
    session <- bow(url, force = TRUE)
    
   top_majors <- scrape(session) |>
     html_nodes(".popular-entity__name") |>
     html_text()
    
    tibble(college = college,
           top_majors = top_majors
    )
}

scrape_majors("university-of-minnesota-twin-cities") #U of M example
```

```{r, echo = FALSE}
college_for_majors <- mn_college_data |>
  select(college_name) |>
  mutate(college = str_to_lower(college_name),
         college = str_replace_all(college, "-", " "),
         college = str_replace_all(college, " ", "-")) |>
  select(-college_name)
```

```{r, echo = FALSE}
#Shiny attempt (the goal is to be able to select a college from a drop down and then get a list of the top 10 majors at that school)

titlePanel("")

inputPanel(
  selectInput("college", label = "college name",
    choices = college_for_majors$college)
)

mainPanel(renderTable({
  scrape_majors <- function(college) {
 
    url <- str_c("https://www.niche.com/colleges/", college, "/majors/")
    
    session <- bow(url, force = TRUE)
    
   top_majors <- scrape(session) |>
     html_nodes(".popular-entity__name") |>
     html_text()
    
   tibble(college = college,
           top_majors = top_majors) |>
     kable()
  }
  
  
}))
```

Recommendations
==========================

Column {data-width = 750}
---------------------------

When it comes to college decisions, there is no one right choice. It depends on what is right for **you**! If you are looking to spend less money, consider a public institution. If you're after smaller class sizes, a private institution might be right for you. Looking for a big school? The U of M is a great option. Spend some time thinking about what is important to you, and go with your gut! 

As three St. Olaf students, our last tip is to check out St. Olaf College! Go Oles! ;)

**Best of luck!**

Column
---------------------------

```{r, echo = FALSE}
renderImage({
    list(src = "images/olaf_grad.png", width = 500, height = 550)
  })
```
