---
title: "MN College Information"
output: 
  flexdashboard::flex_dashboard:
    theme:
      version: 4
      bootswatch: flatly
runtime: shiny
---

About
==========================

Column {data-width = 650}
---------------------------

#### Welcome!

**Hello high-schoolers! As recent high school graduates, we understand that navigating your college decision is difficult. We hope you find this website to be useful as you investigate Minnesota colleges and universities. We have conducted an analysis of multiple factors you might be considering, but please note that everyone cares about different things when evaluating colleges. This is just one place to start! Good luck!**

**-Solveig, Ela, and Morgan**


#### Information

**Authors:**

Solveig Senf, Ela Kanade, & Morgan Ryan

**Data & College Resources**

The data used in this project is from National Center for Education Statistics' IPEDS data base. We compiled data from multiple data sets which all can be accessed [here](https://nces.ed.gov/ipeds/datacenter/DataFiles.aspx?gotoReportId=7&fromIpeds=true&sid=3ce8c2b7-5a48-42a0-baf2-9089ef924924&rtid=7). 

Additionally, we have taken data from ["Niche"](https://www.niche.com/?ref=colleges). It is an excellent resource for information on colleges. Just search for a school!

If you still have unanswered questions about the colleges you are interested in, we recommend visiting [U.S. News Best Colleges](https://www.usnews.com/best-colleges) or spending some time on the college's own website.  

Column
---------------------------

```{r, echo = FALSE}
renderImage({
    list(src = "images/graduate.png", width = 600, height = 600)
  })
```

Minnesota College Map
==========================

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(stringr)
library(rvest)
library(polite)
library(sf)
library(maps)
library(viridis)
library(leaflet)
library(htmltools)
library(readr)
library(ggplot2)
library(ggthemes)
library(ggrepel)
library(plotly)
library(kableExtra)
```

```{r, echo = FALSE, include = FALSE}
#Data Cleaning!
ic <- read_csv("data/ic2023.csv") |>
  select(college_id = UNITID, 
         study_abroad = SLO6, 
         room_capacity = ROOMCAP, 
         ncaa = ASSOC1, 
         college_ran_housing = ROOM, 
         meals_per_week = MEALSWK) 
  
ic_ay <- read_csv("data/ic2023_ay.csv") |>
  select(college_id = UNITID,
         in_state_tuition = TUITION2,
         food_housing_fee = CHG5AY3) 

hd <- read_csv("data/hd2023 2.csv") |>
  select(college_id = UNITID,
         college_name = INSTNM,
         city = CITY,
         state = STABBR,
         open_or_closed = CYACTIVE,
         size_type = INSTSIZE,
         longitude = LONGITUD,
         latitude = LATITUDE,
         public_private = CONTROL)

effy <- read_csv("data/effy2023.csv") |> 
  select(college_id = UNITID,
         student_type = EFFYLEV,
         total_enrollment = EFYTOTLT,
         total_men = EFYTOTLM,
         total_women = EFYTOTLW) |>
  filter(student_type == "2") #filters for undergraduate enrollment numbers 

s <- read_csv("data/s2023_sis.csv") |>
  select(college_id = UNITID,
         employment_type = FACSTAT, 
         total_faculty = SISTOTL) |>
  filter(employment_type == "10") |> #staff with faculty status
  select(-employment_type)
```

```{r, echo = FALSE, include = FALSE}
#Joining clean data sets
ic_full <- ic |>
  left_join(ic_ay)  

ic_hd <- ic_full |>
  left_join(hd)

college_data <- ic_hd |>
  left_join(effy)

college_data2 <- college_data |>
  left_join(s)

mn_college_data <- college_data2 |>
  filter(state == "MN") |>
  arrange(college_name)
```

Column {data-width = 100}
-------------------------

**Wondering about location?**

Check out this map of Minnesota colleges and universities! Click on each marker for the college name, city, and enrollment!

```{r, echo = FALSE}
renderImage({
    list(src = "images/pointing_grad.png", width = 400, height = 450)
  })
```

Column {data-width = 600}
-------------------------

**Minnesota College Map**

```{r, echo = FALSE}

mn_map <- leaflet() |>
  addTiles() |>
  setView(lng = -94.6859, lat = 46.7296, zoom = 6)


mn_map <- mn_map |>
  addMarkers(data = mn_college_data,
             ~longitude, ~latitude,
             popup = ~paste("<b>Name:</b>", college_name, "<br><b>City:</b>", city, "<br><b>Total Enrollment</b>:", total_enrollment))
mn_map
```


Tuition & Type
==========================

An important consideration in your college search is **cost**. College is expensive and it can be really intimidating to decide how much money you are willing to spend. On this page you will find a couple of plots that can help you get a better sense of how much you will be spending depending on the size of your school and whether it is a private or public institution. Note that the tuition expenses are calculated **before** aid, so the cost you would end up paying may be quite different. If you are deciding between a public and private institution, it might be best to apply to both schools, compare prices after aid, and then decide. 

**College by Type**

```{r, echo = FALSE, include = FALSE}
#tibble for tuition by size graph
tuition_by_size <- mn_college_data |>
  select(size_type, in_state_tuition, public_private) |>
  filter(!is.na(in_state_tuition), in_state_tuition != ".") |>
  group_by(size_type, public_private) |>
  mutate(in_state_tuition = as.numeric(in_state_tuition),
         size_type = as.factor(size_type),
         public_private = as.factor(public_private),
         size_type = fct_recode(size_type, 
                                "< 1,000" = "1",
                                "1,000 - 4,999" = "2",
                                "5,000 - 9,999" = "3",
                                "10,000 - 19,999" = "4",
                                "20,000 +" = "5"),
         public_private = fct_recode(public_private,
                                     "Public" = "1",
                                     "Private (not-for-profit)" = "2",
                                     "Private (for-profit)" = "3")) |>
  summarize(mean_tuition = mean(in_state_tuition)) 
```

Column {data-height = 950}
-------------------------

```{r, echo = FALSE}
ui <- fluidPage(
  checkboxInput("check_private", "Show Private Institutions?", value = FALSE),
  tableOutput("college_table")
)

server <- function(input, output) {
  
  output$college_table <- renderTable({
    if (input$check_private) {
      filtered_mn_college_data |>
        filter(`College Type` != "Public")
    } else {
      filtered_mn_college_data |>
        filter(`College Type` == "Public")
    }
  })
}

shinyApp(ui = ui, server = server)
```

```{r, echo = FALSE, include = FALSE}
filtered_mn_college_data <- mn_college_data |>
  select(college_name, public_private) |>
  filter(public_private != "NA") |>
  mutate(public_private = as.factor(public_private),
         public_private = fct_recode(public_private,
                                     "Public" = "1",
                                     "Private (not-for-profit)" = "2",
                                     "Private (for-profit)" = "3")) |>
  rename(college_name, "College Name" = "college_name",
         public_private, "College Type" = "public_private") |>
  arrange(`College Type`) 
  #kable() displays table nicely 
```

Column {data-height = 950}
-------------------------

**Tuition & Size**

```{r, echo = FALSE, warning = FALSE}
inputPanel(
  checkboxInput("type_check", "Color by Institution Type?", value = FALSE)
)
```

```{r, echo = FALSE, warning = FALSE}
renderPlot({
     if (input$type_check) 
        ggplot(tuition_by_size, aes(x = size_type, y = mean_tuition))  + 
          geom_bar(aes(fill = public_private), stat = "identity") +
          theme_classic() +
          scale_fill_colorblind() +
          labs(title = "Average Tuition by Institution Size",
               subtitle = "Colored by Institution Type",
               x = "Size",
               y = "Average Tuition",
               fill = "Type") 
     else if (!input$type_check)      
        ggplot(tuition_by_size, aes(x = size_type, y = mean_tuition))  + 
          geom_bar(stat = "identity") +
          theme_classic() +
          scale_fill_colorblind() +
          labs(title = "Average Tuition by Institution Size",
               subtitle = "Colored by Institution Type",
               x = "Size",
               y = "Average Tuition",
               fill = "Type")
})
```

**Tuition & Type**

```{r, echo = FALSE, warning = FALSE}
mn_college_data |>
  select(in_state_tuition, public_private) |>
  filter(!is.na(in_state_tuition), in_state_tuition != ".") |>
  group_by(public_private) |>
  mutate(in_state_tuition = as.numeric(in_state_tuition),
         public_private = as.factor(public_private),
         public_private = fct_recode(public_private,
                                     "Public" = "1",
                                     "Private (not-for-profit)" = "2",
                                     "Private (for-profit)" = "3")) |>
  summarize(mean_tuition = mean(in_state_tuition)) |>
  ggplot(aes(x = public_private, y = mean_tuition, fill = public_private)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(mean_tuition, 1)), vjust = -0.5, size = 3) +
  labs(title = "Average Tuition by Institution Type",
       x = "Type",
       y = "Average Tuition") +
  theme_classic() +
  scale_fill_colorblind(guide = "none") 
```

Student/Faculty Ratio
==========================

Column {data-height = 1500}
-------------------------

Although it might not seem like it yet, student-to-faculty ratio can play a big role in your college experience. As students who go to a school with a low student-to-faculty ratio, we can attest to how nice it is to have small class sizes and the ability to build relationships with your professors. Professors are a great resource for career exploration, internship searches, and as personal mentors. This plot will help you get a sense of what student-to-faculty ratios look like across Minnesota colleges. 

```{r, echo = FALSE, include = FALSE}
#tibble for student-to-faculty interactive plot
interactive_data <- mn_college_data |>
  select(total_enrollment, total_faculty, college_name, public_private) |>
  filter(total_enrollment != "NA",
         total_faculty != "NA") |>
  mutate(student_to_fac = total_enrollment/total_faculty,
         total_enrollment = as.numeric(total_enrollment),
         public_private = as.factor(public_private),
         public_private = fct_recode(public_private,
                                     "Public" = "1",
                                     "Private (not-for-profit)" = "2",
                                     "Private (for-profit)" = "3")) 
```

```{r, echo = FALSE, include = FALSE}
#tibble for student-to-faculty interactive plot w/out outliers
interactive_data_rm <- mn_college_data |>
  select(total_enrollment, total_faculty, college_name, public_private) |>
  filter(total_enrollment != "NA",
         total_faculty != "NA") |>
  mutate(student_to_fac = total_enrollment/total_faculty,
         total_enrollment = as.numeric(total_enrollment),
         public_private = as.factor(public_private),
         public_private = fct_recode(public_private,
                                     "Public" = "1",
                                     "Private (not-for-profit)" = "2",
                                     "Private (for-profit)" = "3")) |>
  filter(college_name != "University of Minnesota-Twin Cities",
         college_name != "Capella University")
```

```{r, echo = FALSE, warning = FALSE}
inputPanel(
  checkboxInput("type_check", "Remove Outliers?", value = FALSE)
)
```

```{r, echo = FALSE, warning = FALSE}
renderPlotly({
     if (!input$type_check) {
        interactive <- ggplot(interactive_data, aes(x = total_enrollment, y = student_to_fac)) +
        geom_point(aes(text = college_name, color = public_private)) +
        geom_text_repel(aes(label = college_name)) +
        labs(title = "Total Enrollment Vs. Student-to-faculty Ratio",
            x = "Enrollment",
            y = "Student-to-faculty",
            color = "Institution Type") +
        theme_classic() +
        scale_color_colorblind() 
  
        ggplotly(interactive)
     }
     else if (input$type_check) {
        interactive2 <- ggplot(interactive_data_rm, aes(x = total_enrollment, y = student_to_fac)) +
        geom_point(aes(text = college_name, color = public_private)) +
        geom_text_repel(aes(label = college_name)) +
        labs(title = "Total Enrollment Vs. Student-to-faculty Ratio",
            x = "Enrollment",
            y = "Student-to-faculty",
            color = "Institution Type") +
        theme_classic() +
        scale_color_colorblind() 
        
        ggplotly(interactive2)
     }
})
```

Athletics
==========================

Column {data-width = 750}
---------------------------
**Minnesota Intercollegiate Athletic Conference Schools**

This is the place for you if you're a current high school student-athlete looking to play at the next level or an avid sports fan! Below is a list of Division 3 Colleges and Universities in the MIAC (our **favorite** athletic conference in the NCAA). If athletics isn't for you, that's ok too! On the right is a chart showing how many schools participate in the NCAA and how many schools do not. Additionally, we have included a list of schools affiliated with the NCAA. See if the schools you are considering are on the list!

```{r, echo = FALSE, include = FALSE}
# check that scraping is allowed (Step 0)
robotstxt::paths_allowed("https://en.wikipedia.org/wiki/Minnesota_Intercollegiate_Athletic_Conference")

# Step 1: read_html()
miac_colleges <- read_html("https://en.wikipedia.org/wiki/Minnesota_Intercollegiate_Athletic_Conference")

# 2: html_nodes()
miac_tables <- html_nodes(miac_colleges, css = "table") 
miac_tables  # have to guesstimate which table contains climate info

# 3: html_table()
html_table(miac_tables, header = TRUE, fill = TRUE)    # find the right table

miac_data <- html_table(miac_tables, header = TRUE, fill = TRUE)[[2]]  

print(miac_data, n = 10)

library(janitor)

```

```{r, echo = FALSE}
#bow and scrape data
session <- bow("https://en.wikipedia.org/wiki/Minnesota_Intercollegiate_Athletic_Conference", force = TRUE)

miac_result <- scrape(session) |>
  html_nodes(css = "table") |> 
  html_table(header = TRUE, fill = TRUE)

#begin cleaning miac data
miac_data1 <- miac_result[[2]] |>
  clean_names() 

miac_data1 <- miac_data1 |>
  select(institution) |>
  mutate(
    #fixing name consistency
    institution = str_trim(str_replace(institution, "\\[.*\\]", ""), side = "right"),
    institution = str_trim(str_replace(institution, "Mary's University", "Mary's University of Minnesota"), side = "right")
  ) |>
  rename("Institution" = "institution")

miac_data1 |>
  kable()
```

Column
---------------------------
**Colleges and Universities Participating in the NCAA**

```{r, echo = FALSE, include = FALSE}
ncaa <- mn_college_data |>
  select(ncaa) |>
  mutate(ncaa = as.factor(ncaa),
         ncaa = fct_recode(ncaa,
                           "NCAA" = "1",
                           "Not NCAA" = "0",
                           "Unknown" = "-1",
                           "Not Applicable" = "-2")) |>
  count(ncaa)
```

```{r, echo = FALSE}
ncaa |>
  ggplot(aes(x = ncaa, y = n, fill = ncaa)) +
  geom_bar(stat = "identity") +
  labs(title = "NCAA Affiliation",
       x = "Affiliation",
       y = "Count") +
  theme_classic() +
  scale_fill_colorblind(guide = "none") 
```

```{r, echo = FALSE}
ncaa_schools <- mn_college_data |>
  select(college_name, ncaa) |>
  filter(ncaa == "1") |>
  select(-ncaa) |>
  rename("NCAA School" = "college_name")

ncaa_schools |>
  kable()
```


Recommendations
==========================

Column {data-width = 750}
---------------------------

When it comes to college decisions, there is no one right choice. It depends on what is right for **you**! If you are looking to spend less money, consider a public institution. If you're after smaller class sizes, a private institution might be right for you. Looking for a big school? The U of M is a great option. Spend some time thinking about what is important to you, and go with your gut! 

As three St. Olaf students, our last tip is to check out St. Olaf College! Go Oles! ;)

**Best of luck!**

```{r, echo = FALSE}
renderImage({
    list(src = "images/olaf_grad.png", width = 500, height = 550)
  })
```

Column {data-width = 750}
---------------------------

```{r, echo = FALSE, include = FALSE}
robotstxt::paths_allowed("https://www.niche.com/colleges/search/all-colleges/s/minnesota/")

session <- bow("https://www.niche.com/colleges/search/all-colleges/s/minnesota/", force = TRUE)

top_colleges <- scrape(session) |>
  html_nodes(".nss-1s72fo9") |>
  html_text() |>
  as_tibble() |>
  rename(value, "college" = "value")
```

```{r, echo = FALSE}
top_colleges <- top_colleges |>
  mutate(rank = row_number()) |>
  kable()
```
